The genomic impacts of drift and selection for hybrid performance in maize

### 2015-08-27 main.tex
edits to Discussion

Our analysis of the Iowa RRS experiment reveals a steady loss of diversity in the BSSS and BSCB1 populations as they became increasingly differentiated from one another over time.
Principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well. 
These patterns of population structure, diversity, and differentiation between BSSS and BSCB1 can be largely reproduced by simulation without any selection, supporting the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement. 
\rev{Similar observations have been recently made in other reciprocal recurrent selection programs, even in comparisons of multiple replicated populations \citep{romay2012effect,lamkey2014relative}.
These patterns differ markedly, however, from similar experimental evolution in  systems such as \emph{Drosophila} \citep{turner2011population, burke2010genome} and \emph{E. coli} \citep{tenaillon2012molecular} in which effective population sizes are maintained at levels much higher than in the present work and the effects of selection on diversity are readily discernible.}

Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations. 
Numerous experiments have shown that the selected populations and the hybrids formed from them \rev{exhibit genetic gain for hybrid yield as well plant architecture and tolerance to high-density planting} \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population, brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}. 
We find that heterozygosity falls more than expected across the genome as a whole, and, \rev{though drift imposes limitations on the power to detect selection at individual loci,} genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation.  
We further show that an identity-by-descent, haplotype-based approach provides additional power to identify selected regions, as it can distinguish between the fixation of rare and common haplotypes. 
\rev{These analyses show that most likely targets of selection occur at different loci in the two populations, a result consistent with analyses in commercial breeding programs \citep{feng2006temporal} and which may help explain the lack of selection seen in previous analyses across numerous breeding programs \citep{van2012historical}.}

The observation that different targets of selection are  observed in  opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids. 
Classic overdominance models of heterosis predict that at a single locus two distinct alleles confer heterozygote advantage when combined.
Alternatively, the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance). 
In the case of true overdominance, selection should thus lead to decreased heterozygosity at the same locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 
\rev{Although strongly deleterious variants were likely purged during the inbreeding process leading to the founder lines, many weakly deleterious alleles can be found segregating at low frequencies among inbreds \citep{Mezmouk2014}.}
Because deleterious alleles will be rare in both populations, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well.
\rev{Selection against homozygosity of the fixed haplotype will thus} have little impact on diversity in the second population.

Our results are consistent with the dominance and pseudo-overdominance models, but we caution that the exact outcome of particular models will depend strongly on the effects of selection, drift, and the frequencies of beneficial alleles. 
Disentangling these factors remains a major challenge that will require careful simulation of maize breeding using a wider range of parameters. 
This is especially true because in a model of hybrid complementation, genetic drift in one population can alter the selective value of alleles in the other population. 
When drift and selection interact in such a manner, tests for selection that attempt to independently partition the effects of each force may not provide the full picture. 
Furthermore, uncertainties in the fine-scale relationship between genetic and physical distance make it difficult to assign significance in forward-simulation approaches. 
In the end, the best test for selection of specific genomic regions will ultimately be conducted by phenotypic observation in the field in balanced tests of different haplotypes. 
In the case of individual breeding programs, genome-wide genotyping such as we conducted here can identify the lines carrying the recombinant haplotypes and introgressions necessary to conduct such an experiment.

Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}.
Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.  
Nonetheless, breeding has contributed substantially to yield improvements \citep{duvick2005contribution}, and the identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades. 
Several key inbreds in the ‘stiff-stalk’ heterotic group --- B73, B37 and B14 --- were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
%B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.  
Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1. 
Our observation of population-specific selection is consistent with a model of selection for complementation in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm may suggest a role for such selection in maize hybrid improvement in general \citep{van2012historical}.

>>>>

Our analysis of the Iowa RRS experiment reveals a steady loss of diversity in the BSSS and BSCB1 populations as they became increasingly differentiated from one another over time.
Principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well. 
These patterns of population structure, diversity, and differentiation between BSSS and BSCB1 can be largely reproduced by simulation without any selection, supporting the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement. 
\rev{Similar observations have been recently made in other reciprocal recurrent selection programs, even in comparisons of multiple replicated populations \citep{romay2012effect,lamkey2014relative}.
These patterns differ markedly, however, from similar experimental evolution in systems such as \emph{Drosophila} \citep{turner2011population, burke2010genome} and \emph{E. coli} \citep{tenaillon2012molecular} in which the effects of selection on diversity are readily discernible. 
A key difference between these studies and those in maize are the effective population sizes, which were kept much lower to produce a short-term phenotypic response to selection within available field testing resources.} \jpg{My intent in rewording this is to emphasize that Ne is the important difference between the scenarios.}

Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations. 
Numerous experiments have shown that the selected populations and the hybrids formed from them \rev{exhibit genetic gain for hybrid yield as well plant architecture and tolerance to high-density planting} \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population, brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}. 
We find that heterozygosity falls more than expected across the genome as a whole, and, \rev{though drift imposes limitations on the power to detect selection at individual loci,} genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by drift alone.  
We further show that an identity-by-descent, haplotype-based approach provides additional power to identify selected regions, as it can distinguish between the fixation of rare and common haplotypes. 
\rev{These analyses show that the most likely targets of selection occur at different loci in the two populations, a result consistent with analyses in commercial breeding programs \citep{feng2006temporal} and which may help explain the lack of selection seen in previous analyses across numerous breeding programs \citep{van2012historical}.}

The observation that different targets of selection are observed in opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids. 
Classic overdominance models of heterosis predict that at a single locus two distinct alleles confer heterozygote advantage when combined.
Alternatively, the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance). 
In the case of true overdominance, selection should thus lead to decreased heterozygosity at the same locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}. \jpg{One possible criticism is that overdominance could still be happening when each allele is on multiple founder haplotypes. Could say something like "Although we cannot rule out soft sweeps of complementary overdominant alleles, the observed pattern more parsimoniously favors..."}
We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 
\rev{Although strongly deleterious variants were likely purged during the inbreeding process leading to the founder lines, many weakly deleterious alleles can be found segregating at low frequencies among inbreds \citep{Mezmouk2014}.}
Because deleterious alleles will be rare in both populations, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well.
\rev{We expect that selection against homozygosity of the fixed haplotype will thus} have little impact on diversity in the second population.

Although our results better fit the simpler dominance model, we caution that the exact outcome of particular models will depend strongly on the effects of selection, drift, and the frequencies of beneficial alleles. 
Disentangling these factors remains a major challenge that will require careful simulation of maize breeding using a wider range of parameters. 
This is especially true because in a model of hybrid complementation, genetic drift in one population can alter the selective value of alleles in the other population. 
When drift and selection interact in such a manner, tests for selection that attempt to independently partition the effects of each force may not provide the full picture. 
Furthermore, uncertainties in the fine-scale relationship between genetic and physical distance make it difficult to assign significance in forward-simulation approaches. 
In the end, the best test for selection of specific genomic regions will ultimately be conducted by phenotypic observation in the field in balanced tests of different haplotypes. 
In the case of individual breeding programs, genome-wide genotyping such as we conducted here can identify the lines carrying the recombinant haplotypes and introgressions necessary to conduct such an experiment.

Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}.
\jpg{I'm not sure where this paragraph belongs.}Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.  
Nonetheless, breeding has contributed substantially to yield improvements \citep{duvick2005contribution}, and the identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades. 
Several key inbreds in the ‘stiff-stalk’ heterotic group --- B73, B37 and B14 --- were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
%B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.  
Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1. 
Our observation of population-specific selection is consistent with a model of selection for complementation in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm may suggest a role for such selection in maize hybrid improvement in general \citep{van2012historical}.
